---
title: "EDA"
format: html
---

# Prepare

```{r}
#library
library(readr)
library(ggplot2)
library(dplyr)
library(stringr)
library(tidyverse)
library(sf)
library(tidycensus)
```

```{r}
#read CSV
df <- read_csv(
  "D:/MUSA5090/eda/data/opa_properties_public.csv",
  n_max = 100000
)

dim(df)
head(df)
```

```{r}
#Dataset Preview
ggplot(df, aes(x = sale_price)) +
  geom_histogram(bins = 50) +
  labs(
    title = "Sale Price Distribution",
    x = "Sale Price",
    y = "Count"
  )
```

# Data Cleaning and Trimming

```{r}
#Remove non-market & bundle sales

df_clean <- df %>%
  filter(!is.na(sale_price)) %>%
  filter(sale_price > 1)

df_clean <- df_clean %>%
  group_by(sale_date, sale_price) %>%
  filter(n() == 1) %>%
  ungroup()

head(df_clean)
summary(df_clean$sale_price)
```

```{r}
df_clean <- df_clean %>%
  mutate(log_price = log(sale_price))

ggplot(df_clean, aes(x = log_price)) +
  geom_histogram(bins = 50) +
  labs(title = "Log Sale Price Distribution")
```

```{r}
df_residential <- df_clean %>%
  filter(
    str_detect(building_code_description, 
               "ROW|DET|SEMI|S/D|APT|CONDO") &
    !str_detect(building_code_description, 
                "VACANT|COMMERCIAL|IND|STORE|OFF|GAR|HOTEL|AUTO|SCHOOL")
  )

ggplot(df_residential, aes(x = log_price)) +
  geom_histogram(bins = 50) +
  labs(title = "Log Sale Price Distribution")
```

```{r}
#Trim Data
lower <- quantile(df_residential$sale_price, 0.01, na.rm = TRUE)
upper <- quantile(df_residential$sale_price, 0.99, na.rm = TRUE)

lower
upper
```

```{r}
df_final <- df_residential %>%
  filter(sale_price >= lower, sale_price <= upper)

df_final <- df_final %>%
  mutate(log_price = log(sale_price))

summary(df_final$sale_price)
```

To ensure that the analysis reflects actual market behavior, several data cleaning steps were applied to the original dataset.

First, non-market transactions were removed. Observations with extremely low sale prices (e.g., \$1) were excluded, as these likely represent transfers between related parties rather than arms-length transactions. In addition, bundled sales were identified by grouping properties with identical sale_date and sale_price, and observations appearing multiple times within these groups were removed. These transactions likely represent multiple properties sold together and do not reflect individual property values.

Next, the dataset was restricted to residential properties. Given the highly granular building classification system, residential units were identified using keyword-based filtering (e.g., “ROW”, “DET”, “SEMI”, “S/D”, “APT”, “CONDO”), while non-residential categories such as commercial, industrial, and vacant land were excluded. This step ensures that the analysis focuses on comparable property types with similar pricing mechanisms.

Finally, to reduce the influence of extreme values, a two-sided trimming approach was applied using the 1st and 99th percentiles of sale_price. The resulting thresholds were approximately \$3,000 and \$1.7 million. This approach removes anomalous observations while preserving the vast majority of typical market transactions. Although the lower bound remains relatively small, it represents a negligible portion of the data and does not significantly affect the analysis.

# Feature Exploration

## Structural Feartures

```{r}
# Test data missing rate
df <- df %>%
  mutate(log_price = log(sale_price))

structural_vars <- c(
  "total_livable_area",
  "total_area", 
  "number_of_bedrooms",
  "number_of_bathrooms",
  "number_stories",
  "year_built",
  "exterior_condition",
  "interior_condition",
  "quality_grade",
  "type_heater",
  "building_code_description"
)

df %>%
  select(all_of(structural_vars)) %>%
  summarise(across(everything(), ~mean(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_rate") %>%
  arrange(desc(missing_rate))
```

```{r}
#Continuous vs log_price
continuous_vars <- c(
  "total_livable_area",
  "total_area",
  "number_of_bedrooms",
  "number_of_bathrooms",
  "number_stories"
)

df %>%
  select(log_price, all_of(continuous_vars)) %>%
  pivot_longer(-log_price, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = value, y = log_price)) +
  geom_point(alpha = 0.1, size = 0.5) +
  geom_smooth(method = "lm", color = "steelblue", se = FALSE) +
  facet_wrap(~variable, scales = "free_x") +
  labs(title = "Structural continuous features vs log_price")
```

```{r}
#Spearman
df %>%
  select(log_price, all_of(continuous_vars)) %>%
  cor(method = "spearman", use = "complete.obs") %>%
  .[,"log_price"] %>%
  sort(decreasing = TRUE)
```

```{r}
#categorical box plot
categorical_vars <- c(
  "exterior_condition",
  "interior_condition", 
  "quality_grade",
  "building_code_description"
)

df %>%
  select(log_price, all_of(categorical_vars)) %>%
  mutate(across(all_of(categorical_vars), as.character)) %>%  
  pivot_longer(-log_price, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = as.factor(value), y = log_price)) +
  geom_boxplot(outlier.alpha = 0.1) +
  facet_wrap(~variable, scales = "free_x") +
  labs(title = "Structural categorical features vs log_price") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```


```{r}
#Total livable area
df <- df %>%
  filter(is.na(total_livable_area) | total_livable_area < 10000)

df <- df %>%
  mutate(quality_grade_simplified = str_extract(quality_grade, "^[A-Z]"))

```

```{r}
#correlation
cor(df$total_livable_area, df$log_price, 
    method = "spearman", use = "complete.obs")
```

```{r}
#Quality grade
df %>%
  filter(!is.na(quality_grade_simplified)) %>%
  ggplot(aes(x = quality_grade_simplified, y = log_price)) +
  geom_boxplot(outlier.alpha = 0.1) +
  labs(title = "quality_grade_simplified vs log_price")
```

```{r}
#Age
df <- df %>%
  mutate(age = 2025 - year_built,
         age = ifelse(age < 0 | age > 200, NA, age))

cor(df$age, df$log_price, 
    method = "spearman", use = "complete.obs")
```

| feature                     | 保留？               |
|-----------------------------|----------------------|
| number_of_bathrooms         | 保留，cap 异常值     |
| total_livable_area          | 保留                 |
| exterior_condition          | 保留                 |
| interior_condition          | 保留，之后查共线性   |
| quality_grade_simplified    | 保留，C/D 噪音大     |
| age                         | 保留，信号弱但方向对 |
| number_of_bedrooms          | 可选，信号弱         |
| number_stories / total_area | 丢掉                 |
| type_heater                 | 丢掉                 |

## Location Features

```{r}
# missing rate
df %>%
  select(zip_code, census_tract, geographic_ward, zoning) %>%
  summarise(across(everything(), ~mean(is.na(.)))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_rate")

# zip_code
df %>%
  group_by(zip_code) %>%
  summarise(
    median_log_price = median(log_price, na.rm = TRUE),
    n = n()
  ) %>%
  filter(n > 50) %>%
  arrange(desc(median_log_price))
```

```{r}
# geographic_ward
df %>%
  group_by(geographic_ward) %>%
  summarise(
    median_log_price = median(log_price, na.rm = TRUE),
    n = n()
  ) %>%
  filter(n > 50) %>%
  arrange(desc(median_log_price)) %>%
  head(10)

# census_tract
df %>%
  summarise(
    n_zip = n_distinct(zip_code),
    n_ward = n_distinct(geographic_ward),
    n_tract = n_distinct(census_tract)
  )
```

```{r}
df %>%
  group_by(census_tract) %>%
  summarise(
    median_log_price = median(log_price, na.rm = TRUE),
    n = n()
  ) %>%
  filter(n > 30) %>%
  summarise(
    tract_price_sd = sd(median_log_price),
    zip_reference = "compare with zip"
  )

df %>%
  group_by(zip_code) %>%
  summarise(median_log_price = median(log_price, na.rm = TRUE)) %>%
  summarise(zip_price_sd = sd(median_log_price))
```

```{r}
df %>%
  select(log_price, topography, view_type, site_type) %>%
  mutate(across(-log_price, as.character)) %>%
  pivot_longer(-log_price, names_to = "variable", values_to = "value") %>%
  ggplot(aes(x = as.factor(value), y = log_price)) +
  geom_boxplot(outlier.alpha = 0.1) +
  facet_wrap(~variable, scales = "free_x") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# missing rate
df %>%
  summarise(
    missing_rate = mean(is.na(zoning)),
    n_distinct = n_distinct(zoning, na.rm = TRUE)
  )

# signal
df %>%
  group_by(zoning) %>%
  summarise(
    median_log_price = median(log_price, na.rm = TRUE),
    n = n()
  ) %>%
  filter(n > 50) %>%
  arrange(desc(median_log_price))
```

```{r}
df %>%
  summarise(
    missing_rate = mean(is.na(zoning)),
    n_distinct = n_distinct(zoning, na.rm = TRUE)
  )

df %>%
  count(zoning, sort = TRUE) %>%
  summarise(n_categories = n())
```

```{r}
df <- df %>%
  mutate(zoning_prefix = str_extract(zoning, "^[A-Z]+"))

df %>%
  count(zoning_prefix, sort = TRUE)

df <- df %>%
  mutate(zoning_prefix = case_when(
    zoning_prefix %in% c("RSA", "RM", "CMX", "RSD", 
                          "RMX", "RTA", "ICMX") ~ zoning_prefix,
    is.na(zoning_prefix) ~ NA_character_,
    TRUE ~ "other"
  ))
```

| eature                                   | 保留？                |
|------------------------------------------|-----------------------|
| `zip_code`                               | 保留                  |
| `zoning_prefix`                          | 保留，简化自 `zoning` |
| `census_tract`                           | 丢掉                  |
| `geographic_ward`                        | 丢掉                  |
| `topography` / `view_type` / `site_type` | 全部丢掉              |

## Temproal Features

```{r}
df <- df %>%
  mutate(
    sale_year = year(sale_date),
    sale_month = month(sale_date),
    sale_quarter = quarter(sale_date),
    sale_age = 2025 - sale_year
  )

# seasonal
df %>%
  group_by(sale_month) %>%
  summarise(
    median_log_price = median(log_price, na.rm = TRUE),
    n = n()
  ) %>%
  ggplot(aes(x = sale_month, y = median_log_price)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 1:12) +
  labs(title = "Median log_price by sale month")
```

```{r}
# sale volume
df %>%
  count(sale_year) %>%
  filter(!is.na(sale_year)) %>%
  ggplot(aes(x = sale_year, y = n)) +
  geom_col() +
  labs(title = "Sale volume by year")

# sale year
df %>%
  group_by(sale_year) %>%
  summarise(median_log_price = median(log_price, na.rm = TRUE)) %>%
  filter(!is.na(sale_year)) %>%
  ggplot(aes(x = sale_year, y = median_log_price)) +
  geom_line() + geom_point() +
  labs(title = "Median log_price by sale year")
```

```{r}
df <- df %>%
  filter(sale_year >= 2021)

df %>% count(sale_year)
```

| feature                              | 保留？                    |
|--------------------------------------|---------------------------|
| `sale_year`                          | 保留                      |
| `sale_month`                         | 保留，季节性信号强        |
| `sale_age`                           | 丢掉，数据只有3-4年没意义 |
| `age`（year_built算的）              | 保留，信号弱但方向对      |
| `recording_date` / `assessment_date` | 丢掉，和sale_date重复     |

## External Features

```{r}
crime <- read_csv("data/crime.csv")
ppr <- read_csv("data/Recreation.csv")
transit <- read_csv("data/transit.csv")

glimpse(crime)
glimpse(ppr)
glimpse(transit)
```

```{r}
df_sf <- df %>%
  filter(!is.na(shape)) %>%
  st_as_sf(wkt = "shape", crs = 2272) %>%  
  st_transform(crs = 4326)                 

df <- df_sf %>%
  mutate(
    lng = st_coordinates(.)[,1],
    lat = st_coordinates(.)[,2]
  ) %>%
  st_drop_geometry()                      

df %>% select(lat, lng) %>% head()
```

**Spatial join**

```{r}
crime_sf <- crime %>%
  filter(!is.na(lat), !is.na(lng)) %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326)

opa_sf <- df %>%
  filter(!is.na(lat), !is.na(lng)) %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326) %>%
  st_transform(crs = 4326)

opa_buffer <- opa_sf %>%
  st_transform(crs = 3857) %>%   
  st_buffer(500)                 # 500m buffer

crime_sf_meter <- crime_sf %>%
  st_transform(crs = 3857)

# spatial join + count
crime_count <- st_join(opa_buffer, crime_sf_meter) %>%
  st_drop_geometry() %>%
  group_by(parcel_number) %>%
  summarise(crime_count_500m = n())

# join df
df <- df %>%
  left_join(crime_count, by = "parcel_number")
```

```{r}
summary(df$crime_count_500m)
```

```{r}
cor(df$crime_count_500m, df$log_price, 
    method = "spearman", use = "complete.obs")
```

```{r}
ppr_sf <- ppr %>%
  filter(!is.na(X), !is.na(Y)) %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326) %>%
  st_transform(crs = 3857)

opa_meter <- opa_sf %>%
  st_transform(crs = 3857)

# parcel to park dist
nearest_park <- st_nearest_feature(opa_meter, ppr_sf)
df$dist_to_park <- st_distance(
  opa_meter, 
  ppr_sf[nearest_park,], 
  by_element = TRUE
) %>% as.numeric()

summary(df$dist_to_park)
```

```{r}
cor(df$dist_to_park, df$log_price,
    method = "spearman", use = "complete.obs")
```

```{r}
transit_sf <- transit %>%
  filter(!is.na(Lat), !is.na(Lon)) %>%
  st_as_sf(coords = c("Lon", "Lat"), crs = 4326) %>%
  st_transform(crs = 3857)

nearest_stop <- st_nearest_feature(opa_meter, transit_sf)
df$dist_to_transit <- st_distance(
  opa_meter,
  transit_sf[nearest_stop,],
  by_element = TRUE
) %>% as.numeric()

summary(df$dist_to_transit)
```

```{r}
cor(df$dist_to_transit, df$log_price,
    method = "spearman", use = "complete.obs")
```

```{r}
library(sf)
library(tidyverse)

hospital <- st_read("data/Hospitals.geojson")
school <- st_read("data/Schools_Parcels.geojson")

glimpse(hospital)
glimpse(school)
```
```{r}
# only active school data
school_active <- school %>%
  filter(ACTIVE == "O" | is.na(ACTIVE) == FALSE)

# TYPE_SPECIFIC 
school %>% count(TYPE_SPECIFIC)

# GRADE_LEVEL  
school %>% count(GRADE_LEVEL)
```
```{r}
opa_sf <- df %>%
  filter(!is.na(lat), !is.na(lng)) %>%
  st_as_sf(coords = c("lng", "lat"), crs = 4326)

opa_meter <- opa_sf %>%
  st_transform(crs = 3857)

#only public
school_public <- school %>%
  filter(TYPE_SPECIFIC %in% c("DISTRICT", "CHARTER"))

school_meter <- school_public %>%
  st_transform(crs = 3857) %>%
  st_centroid()  

opa_meter <- opa_sf %>%
  st_transform(crs = 3857)

nearest_school <- st_nearest_feature(opa_meter, school_meter)
df$dist_to_school <- st_distance(
  opa_meter,
  school_meter[nearest_school,],
  by_element = TRUE
) %>% as.numeric()

#Hospital
hospital_meter <- hospital %>%
  st_transform(crs = 3857)

nearest_hospital <- st_nearest_feature(opa_meter, hospital_meter)
df$dist_to_hospital <- st_distance(
  opa_meter,
  hospital_meter[nearest_hospital,],
  by_element = TRUE
) %>% as.numeric()
```

```{r}
cor(df %>% select(log_price, dist_to_school, dist_to_hospital),
    method = "spearman", use = "complete.obs")
```


| feature            | 相关系数 | 保留？ |
|--------------------|----------|--------|
| `crime_count_500m` | -0.17    | 保留   |
| `dist_to_park`     | 0.04     | 丢掉   |
| `dist_to_transit`  | -0.03    | 丢掉   |

### **Census**

```{r}
census_data <- get_acs(
  geography = "tract",
  state = "PA",
  county = "Philadelphia",
  variables = c(
    median_income    = "B19013_001",
    poverty_rate     = "B17001_002",
    bach_degree_rate = "B15003_022"
  ),
  year = 2022,
  geometry = TRUE  
)
```

```{r}
census_sf <- census_data %>%
  st_transform(crs = 3857)

opa_meter <- opa_sf %>%
  st_transform(crs = 3857)

df_census <- st_join(opa_meter, census_sf) %>%
  st_drop_geometry() %>%
  select(parcel_number, variable, estimate) %>%
  pivot_wider(names_from = variable, values_from = estimate)

df <- df %>%
  left_join(df_census, by = "parcel_number")
```

```{r}
cor(df %>% select(log_price, median_income, poverty_rate, bach_degree_rate),
    method = "spearman", use = "complete.obs")
```

| feature            | 相关系数 | 方向                       |
|--------------------|----------|----------------------------|
| `median_income`    | 0.41     | 正确，收入越高房价越高     |
| `bach_degree_rate` | 0.36     | 正确，教育水平越高房价越高 |
| `poverty_rate`     | -0.35    | 正确，贫困率越高房价越低   |

**Interaction**

```{r}
# zip_code x total_livable_area
df %>%
  filter(!is.na(zip_code), !is.na(total_livable_area)) %>%
  group_by(zip_code) %>%
  mutate(area_quintile = ntile(total_livable_area, 5)) %>%
  group_by(zip_code, area_quintile) %>%
  summarise(median_log_price = median(log_price, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = area_quintile, y = median_log_price, 
             group = zip_code, color = zip_code)) +
  geom_line(alpha = 0.5) +
  labs(title = "log_price vs area quintile by zip_code",
       x = "area quintile (1=smallest, 5=largest)",
       y = "median log_price") +
  theme(legend.position = "none")
```
